---
title: "EDA"
output: html_document
---


```{r}
rsconnect::setAccountInfo(name=)
```


```{r}


library("readr")

r=read.csv("winequality-red.csv", sep=";", header=TRUE)
w=read.csv("winequality-white.csv", sep=";", header=TRUE)
r
w

ggplot(w, aes(x=quality, y=density)) + geom_point()



names(w)

```

```{r}

#install.packages("GGally")
library("GGally")
library("ggplot2")
ggpairs(r)




```

```{r}

ggpairs(w)

```

Correlation matrix:

```{r}
#install.packages("qtlcharts")

library(qtlcharts)
#note: for the ones greater than 0.2
iplotCorr(r) # citric acid, sulphates, alcohol | acidity (neg)
iplotCorr(w)#alcohol | chlorides and density (neg)



```



```{r}
library(ggplot2)
library(ggfortify)

```

```{r}


summary(r)
ggplot(r, aes(x=quality))+geom_histogram(bins = 6)

summary(w)
ggplot(w, aes(x=quality))+geom_histogram(bins = 6)

```
```{r}
r
```


```{r}


# citric acid, sulphates, alcohol | acidity (neg)
library(dplyr)

r2=r%>%select(volatile.acidity,sulphates,alcohol,quality)
ggpairs(r2)

w2 =w%>% select(quality, alcohol, volatile.acidity, residual.sugar, free.sulfur.dioxide, density, pH, sulphates, fixed.acidity)
ggpairs(w2)

```

Checking for linearity + assumptions = get rid pof sulphates
```{r}

ggplot(r2, aes(x=sulphates,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=F)
ggplot(r2, aes(x=volatile.acidity,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=F)
ggplot(r2, aes(x=alcohol,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=F)




lms=lm(quality~sulphates, data=r2)
autoplot(lms, which =c(1,2))

lmv=lm(quality~volatile.acidity, data=r2)
autoplot(lmv, which =c(1,2))

lma=lm(quality~alcohol, data=r2)
autoplot(lma, which =c(1,2))


lmt=lm(quality~volatile.acidity+alcohol, data=r2)
autoplot(lmt, which =c(1,2))
summary(lmt)

```
```{r}
lm1=lm(quality~alcohol+sulphates+volatile.acidity+residual.sugar+pH+fixed.acidity,data=r)
lm1
```

```{r}
ggplot(w2, aes(x=alcohol,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=F)
lms=lm(quality~sulphates, data=w2)
autoplot(lms, which =c(1,2))
lms=lm(quality~alcohol, data=w2)
autoplot(lms, which =c(1,2))
lms=lm(quality~volatile.acidity, data=w2)
autoplot(lms, which =c(1,2))
lms=lm(quality~residual.sugar, data=w2)
autoplot(lms, which =c(1,2))
lms=lm(quality~free.sulfur.dioxide, data=w2)
autoplot(lms, which =c(1,2)) #No dice
lms=lm(quality~density, data=w2)
autoplot(lms, which =c(1,2)) # do dice
lms=lm(quality~pH, data=w2)
autoplot(lms, which =c(1,2))
lms=lm(quality~fixed.acidity, data=w2)
autoplot(lms, which =c(1,2))

#(quality, alcohol, volatile.acidity, residual.sugar, free.sulfur.dioxide, density, pH, sulphates, fixed.acidity)
```

```{r}


lm1=lm(quality~.,data=r)
autoplot(lm1,which=c(1:2))
lmb=step(lm1)
autoplot(lmb,which=c(1:2))
summary(lmb)
```
```{r}

lm0=lm(quality~1,data=r)
lmf=step(lm0,scope=list(lower=lm0,upper=lm1),direction="forward")
autoplot(lmf,which=c(1:2))
summary(lmf)
```
install.packages("leaps")
```{r}
lm1=lm(quality~.,data=w)
autoplot(lm1,which=c(1:2))
lmb=step(lm1)
autoplot(lmb,which=c(1:2))
summary(lmb)
```
```{r}

lm0=lm(quality~1,data=w)
lmf=step(lm0,scope=list(lower=lm0,upper=lm1),direction="forward")
autoplot(lmf,which=c(1:2))
summary(lmf)
```

```{r}

library(leaps)
exh=regsubsets(quality~.,data=r)
summary(exh)$outmat

```

get rid of :
free.sulfur.dioxide --> not linear+norm
total sf - notlinear+norm
```{r}
exh=regsubsets(quality~.,data=r)
summary(exh)$outmat
```

bakward lm model

Call:
lm(formula = quality ~ alcohol + volatile.acidity + residual.sugar + 
    free.sulfur.dioxide + density + pH + sulphates + fixed.acidity, 
    data = w)
    
    
    
    
    
    
    
    
    
    
      WHITE WINEMODEL - RECHCKING LINEARITY 
    
    
    
    
    


```{r}


ggplot(w, aes(x=alcohol,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=volatile.acidity,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=residual.sugar,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=free.sulfur.dioxide,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=density,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)

ggplot(w, aes(x=pH,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=sulphates,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)
ggplot(w, aes(x=fixed.acidity,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)

ggplot(w, aes(x=chlorides,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


```
white wine - get rid of sulphates,ph, free sulfur dioxide, resdual sugar, 

FINAL MODEL:Alcohol,olatile.acidity, density, fixed acidiyy

```{r}
w4=w%>%select(quality,alcohol,volatile.acidity, density,  fixed.acidity)
vif(w4)

```


redoing stepwise wiht the following airbales

```{r}

wl1=lm(quality~.,data=w4)
wl0=lm(quality~1,data=w4)
wlf=step(wl0,scope=list(lower=wl0,upper=wl1),direction="forward")
autoplot(wlf,which=c(1:2))
summary(wlf)

wlb=step(wl1)
autoplot(wlb)
summary(wlb)

tab_model(wlb, wlf,show.aic=T,dv.labels=c("Backward stepwise for White Wine","Forward stepwise for White Wine"))



```






























```{r}

lcheck=lm(quality~free.sulfur.dioxide, data=r)
autoplot(lcheck, which =c(1,2))

ggplot(r, aes(x=free.sulfur.dioxide,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


lmvolatile.acidity=lm(quality~volatile.acidity, data=r)
autoplot(lmvolatile.acidity, which =c(1,2))

ggplot(r, aes(x=volatile.acidity,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


lmtotal.sulfur.dioxide=lm(quality~total.sulfur.dioxide, data=r)
autoplot(lmtotal.sulfur.dioxide, which =c(1,2))

ggplot(r, aes(x=total.sulfur.dioxide,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


lmtotal.sulfur.dioxide=lm(quality~total.sulfur.dioxide, data=r)
autoplot(lmtotal.sulfur.dioxide, which =c(1,2))

ggplot(r, aes(x=total.sulfur.dioxide,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


lmchlorides=lm(quality~chlorides, data=r)
autoplot(lmchlorides, which =c(1,2))

ggplot(r, aes(x=chlorides,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


ggplot(r, aes(x=pH,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)

ggplot(r, aes(x=free.sulfur.dioxide,y=quality))+geom_point(size=2)+theme_classic()+geom_smooth(method="lm", se=T)


```














```{r}
library(sjPlot)
sjPlot::tab_model(lmf,lmb,show.ci=T,show.aic=T,dv.labels=c("Forward model","Backward model"))



```




redo

```{r}

r4=r%>%select(quality,alcohol,sulphates,volatile.acidity)

lm1=lm(quality~.,data=r4)
lmbf=step(lm1)
autoplot(lmbf,which=c(1:2))
summary(lmbf)

lm0=lm(quality~1,data=r4)
lmff=step(lm0,scope=list(lower=lm0,upper=lm1),direction="forward")
autoplot(lmff,which=c(1:2))
summary(lmff)


sjPlot::tab_model(lmbf,lmff,show.aic=T,dv.labels=c("Backward stepwise for Red Wine","Forward stepwise for Red Wine"))
```
```{r,fig.width=6,fig.height=2 }
autoplot(lmbf,which=c(1:2))
autoplot(wlf,which=c(1:2))

```
```{r, results = "asis"}
equatiomatic::extract_eq(lmbf, intercept = "beta",use_coefs = TRUE)

equatiomatic::extract_eq(wlf, intercept = "beta",use_coefs = TRUE)



```

$$
\operatorname{Quality} = 2.61 + 0.31(\operatorname{alcohol}) + 0.68(\operatorname{sulphates}) - 1.22(\operatorname{volatile acidity}) + \epsilon
$$$$



\operatorname{Quality} = -48.57 + 0.41(\operatorname{alcohol}) - 2.11(\operatorname{volatile acidity}) + 51.67(\operatorname{density}) - 0.1(\operatorname{fixed acidity}) + \epsilon
$$



```{r}
install.packages("lava",type="binary")
library(caret)
```


```{r}

cross=train(quality~alcohol+sulphates+volatile.acidity,data=r, method="lm",trControl=trainControl(method="cv",number=10,verboseIter=FALSE))

cross



```




```{r}
crossw=train(quality~alcohol+volatile.acidity+ density+fixed.acidity+chlorides , data=w, method="lm",trControl=trainControl(method="cv",number=10,verboseIter=FALSE))

crossw

#(quality, alcohol, volatile.acidity, residual.sugar, pH, sulphates, fixed.acidity)
```

```{r}
knitr::kable(crossw$results,align="cll",dig=2,caption = "Predictive performance of multiple linear regression for White wine quality predictions.") %>% kable_paper(full_width = F,html_font = "Cambria")
```

```{r}
knitr::kable(cross$results,align="cll",dig=2,caption = "Predictive performance of multiple linear regression for Red wine quality predictions.") %>% kable_paper(full_width = F,html_font = "Cambria")
```




```{r}

modelinsampler=r%>%mutate(predicted=round(predict(lmbf,r)))%>% select(quality, predicted)
modelinsampler
insampler=mean(modelinsampler$quality==modelinsampler$predicted)
insampler

```
```{r}

modelinsamplew=w%>%mutate(predicted=round(predict(wlb,white)))%>% select(quality, predicted)
modelinsamplew
insamplew=mean(modelinsamplew$quality==modelinsamplew$predicted)
insamplew


```
```{r}
plot_model(lmff, type="pred", terms=c("sulphates", "alcohol","volatile.acidity"))
```

